Formal Analysis of EDHOC Key Establishment for Constrained IoT Devices

### 2022-07-15 paper.tex
Updates from Overleaf

%-------------------------------------------------------------------------- sec
\section{\uppercase{Formalization and Results}}
\label{sec:formalization}
\input{formalization}

@@

%-------------------------------------------------------------------------- sec
\section{\uppercase{Error handling}}
\label{sec:errorHandling}
\mPoint{Error handling is undefined except for algo negotiation}
Section 6 of the \mEdhoc{} specification\footnote{draft-ietf-lake-edhoc-14}
states that error messages can be sent at any time and by any party.
%
One error message type is defined (for algorithm negotiation), and the rest can
be arbitrarily defined by the implementation.
%
The contents and semantics of these messages are also implementation defined.
%
Consequently, the error messages provide a generic message passing mechanism.
%

\textbf{Q:} Is htere any situation where it makes sense that the initiator sends
a unsupported ciphersuite message instead of message 3?
%

\textbf{Q:} In Section H: if no message correlation support by the transport
layer, \mCi{} and \mCr{} should be used for correlation of messages to a
session.
%
Error messages do not carry \mCi{} or \mCr{}; will this create a problem?
%

\mEdhoc{} provides the error message mechanism, but gives little to no guidance
on how it should be used safely.
%
Example of issue with improper use: Assume application logs error messages a
finite log with log rotation.
%
If the log is used for anomaly detection or detection of sensitive events, then
an adversary can simply inject error messages and fill the log until the
sensitive event is overwritten.
%
If the filling error messages are less suspicious than the actual attack, this
may be beneficial for the adversary.
%
Implementors should do proper log-separation and log management.
%

Section 6.2 gives an excample of sending an error message "method not
supported".
%
This message cannot be used to negotiate method securely.
%
The reason is that error messages are not integrity protected, and an initator
only proposes one method, and therefore the same tactic used for ciphersuite
negotiation cannot be used for methods.
%
Because method negotiation can be viewed in the context of crypto agility, there
is room for improvement.
%

>>>>

%-------------------------------------------------------------------------- sec
\section{\uppercase{Formalization and Results}}
\label{sec:formalization}
\input{edhocFormalization}

@@

%-------------------------------------------------------------------------- sec
\section{\uppercase{Error handling}}
\label{sec:errorHandling}
\mPoint{Error handling is undefined except for algo negotiation} Section 6 of the \mEdhoc{} specification\cite{draft-ietf-lake-edhoc-15} states that error messages can be sent at any time and by any party.
%
At the current version of the standard \cite{draft-ietf-lake-edhoc-15}, there are three types of error messages: type 0 is used to represent success, type 1 represents a generic error message used for debugging purposes, where the actual message is an English string, and type 2 represents failure to negotiate a common ciphersuite. Type 2 messages carry a list of supported ciphersuites by the responder.
%
The contents and semantics of type 1 messages are dependent on the implementation.
%
Consequently, type 1 error messages provide a generic message passing mechanism, albeit one without a predefined semantics.

%\textbf{Q:} Is there any situation where it makes sense that the initiator sends an unsupported ciphersuite message instead of message 3?
%\textbf{Q:} In Section H: if no message correlation support by the transport layer, \mCi{} and \mCr{} should be used for correlation of messages to a session.

\subsection{Lack of connection identifiers in error messages}
Error messages do not include the optional connection identifiers \mCi{} or \mCr{} that are transmitted during the handshake.
This is potentially problematic for connection-less transport layers, where \mCi{} and \mCr{} may be used to correlate messages.

\subsection{Improper handling of error messages}
\mEdhoc{} provides the error message mechanism, but gives little to no guidance on how it should be used safely.
%
For example, assume that an application using \mEdhoc{} logs error messages on a finite log with log rotation.
%
If the log is used for anomaly detection or detection of sensitive events, then an adversary can simply inject error messages and fill the log until the sensitive event is overwritten.
%
If filling error messages in the log is less suspicious than the actual attack, this may be beneficial for the adversary.
%
Implementors should do proper log-separation and log management.

\subsection{Unspecified semantics for debugging messages}
Section 6.2 gives an example of sending a type 1 error message with the information string "Method not supported".
%
It's important to note that this message cannot be used to negotiate the connection method securely.
%
In fact -- differently from ciphersuite negotiation -- \mEdhoc{} does not include a mechanism to support negotiating the connection method, among the four available combinations of signature keys and static Diffie-Hellman keys.
%
The initiator selects the method on the first message, and the responder can reply with a type 1 error message along with the information string "Method not supported", but a standard compliant implementation cannot act upon this information string for negotiation purposes.
%
This is because the error message is not integrity protected and there is no secure mechanism to select a different authentication method that ensure that the preferred available authentication mechanism is always selected.
%
Because method negotiation can be viewed in the context of cryptographic agility, there is room for improvement.
%
